The Sound of Teaching Music: Expert Pianists' Performance Modulations for Novices

### 2020-12-15 paper/manuscript.tex
some corrections for plots
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics). A paired-sample \emph{t}-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition. For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively). The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA. For post-hoc comparisons, we used emmeans package to investigate interactions from the ANOVA.

@@

\hypertarget{kv-transition-points}{%
\subsubsection{2.2.4. KV Transition Points}\label{kv-transition-points}}

@@

\hypertarget{kv-transition-points-1}{%
\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points-1}}

@@

@@

\hypertarget{kv-transition-points-2}{%
\subsubsection{3.2.4. KV Transition Points}\label{kv-transition-points-2}}

@@

\hypertarget{kv-transition-points-3}{%
\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-3}}

@@

>>>>

We performed separate analyses for the two expressive techniques (i.e., articulation, dynamics). A paired-sample \emph{t}-test or a Wilcoxon signed-rank tests (a non-parametric alternative to a paired \emph{t}-test) was performed to compare the mean IOIs in the teaching and performing condition. For KOT and KV, we performed a 2 x 2 repeated-measures analysis of variance (ANOVA) with the factors Condition (teaching vs.~performing) and Subcomponent (LS; legato vs.~staccato or FP; forte vs.~piano, respectively). The ``aov\_car" function in the afex package was used for a repeated-measures ANOVA. For post-hoc comparisons, we used emmeans package.

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant interaction between Condition and Subcomponent-LS (\emph{F}(1,30) = 7.38, \emph{p} = 0.011, \(\eta_G^2\) = 0.0044). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not show a significant difference between the teaching and performing condition for each subcomponent (Legato; \emph{p} = 0.696, Staccato; \emph{p} = 0.07).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 369.54, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7367) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,30) = 6.80, \emph{p} = 0.014, \(\eta_G^2\) = 0.0034). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte in the teaching condition than in the performing condition (\emph{p} = 0.021) while there was no significant difference in piano between the teaching and performing condition (\emph{p} = 0.448).

\hypertarget{kv-transition-points}{%
\subsubsection{2.3.3. KV Transition Points}\label{kv-transition-points}}

Additionally, we conducted an exploratory analysis by focusing on specific points at which each subcomponent changed from one to the other (i.e., forte to piano; FtoP, piano to forte; PtoF). These points could be structurally important in order to make a contrast between forte and piano. We calculated a KV difference for each interval subtracting the KV value of the current note from that of the following note. A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,30) = 223.89, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8344) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,30) = 25.74, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.0321). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.002, PtoF; \emph{p} \textless{} 0.001.

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_text).
\end{verbatim}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-1-1} \caption{\label{fig:vel-1}Experiment 1: KV (0-127) when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-1}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-diff-1-1} \caption{\label{fig:vel-diff-1}Experiment 1: KV Difference at transition points when playing the piece with dynamics. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-diff-1}
\end{figure}

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-LS (legato vs.~staccato) showed that there was a significant main effect of Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.011, \(\eta_G^2\) = 0.0400), reflecting louder sound for legato notes compared to staccato notes. Neither the main effect of Condition (\emph{F}(1,18) = 0.46, \emph{p} = 0.506, \(\eta_G^2\) = 0.0008) nor the interaction between Condition and Subcomponent-LS (\emph{F}(1,18) = 8.15, \emph{p} = 0.799, \(\eta_G^2\) = 0.0001).

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (forte vs.~piano) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 126.10, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.7190) and a significant interaction between Condition and Subcomponent-FP (\emph{F}(1,18) = 6.11, \emph{p} = 0.024, \(\eta_G^2\) = 0.0217). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that participants produced louder forte and softer piano in the teaching condition than in the performing condition (Forte; \emph{p} = 0.036, Piano; \emph{p} = 0.028).

\hypertarget{kv-transition-points-1}{%
\subsubsection{3.3.3. KV Transition Points}\label{kv-transition-points-1}}

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Subcomponent-FP (FtoP vs.~PtoF) showed that there was a significant main effect of Subcomponent-FP (\emph{F}(1,18) = 110.64, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.8167) and a significant interaction of Condition and Subcomponent-FP (\emph{F}(1,18) = 9.77, \emph{p} = 0.006, \(\eta_G^2\) = 0.0641). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment showed that there was a larger KV difference when the expressive notation changed from forte to piano and from piano to forte in the teaching condition than in the performing condition (FtoP; \emph{p} = 0.017, PtoF; \emph{p} = 0.003.

\begin{verbatim}
## Warning: Removed 2 rows containing missing values (geom_text).
\end{verbatim}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-vel-2-1} \caption{\label{fig:vel-2}Experiment 2: KV (0-127) when playing the piece with either articulation (left) or dynamics (right). Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box. Significance levels: * \textit{p} < .05, ** \textit{p} < .01, *** \textit{p} < .001}\label{fig:plot-vel-2}
\end{figure}
